---
title: "Learning RMarkdown"
author: "Gergely"
date: "February 4, 2017"
output: html_document
---


Markdown is a simple markup language to format plain text files, eg using *italic* or **bold** font, or sometimes even ***both*** or ~~none~~ at all. We can also refer to inline `code` examples, or use syntax highlighting for larger code chunks:

```r
n <- 100
x <- rpois(n, 10)
summary(x)
hist(x)
```

But we can also evaluate `r "R"` code inline or in separate code chunks as well:

```{r}
n <- 100
x <- runif(n)
summary(x)
hist(x)
## TODO show error
```


We can generate markdown tables to be rendered nicely in PDF/HTML/docx etc documents by `kable` from the `knitr` package, or `pander` for the more complex tables:

```{r error=TRUE}
knitr::kable(head(iris))

library(pander)
pander(head(iris))

knitr::kable(table(iris$Species))
pander(table(iris$Species))
```


And this latter `pander` package has some nifty features:

```{r}
fit <- lm(mpg ~ wt + hp, mtcars)
fit
pander(fit)

emphasize.strong.cells(which(head(mtcars) < 5, arr.ind = TRUE))
pander(head(mtcars))
```

We can suppress the R commands easily with one of the many chunk options:

```{r, echo = FALSE}
pander(table(iris$Species), justify = 'right')
```

Or we can use global options, eg:

```{r global_options, include = FALSE}
library(knitr)
opts_chunk$set(
    fig.width  = 8,
    fig.height = 8,
    dpi        = 300,
    echo       = FALSE,
    warning    = FALSE,
    message    = FALSE)
```

Dynamic stuff:

```{r}
library(shiny)
inputPanel(
    sliderInput('n', label = 'N', min = 1, max = 1000, value = 100, step = 10),
    sliderInput('seed', label = 'Seed', min = 1, max = 42, value = 42, step = 1)) ## LATER

renderPlot({
  set.seed(as.numeric(input$seed)) ## LATER
  hist(runif(as.numeric(input$n)))
})
```

So now you have everything in hand to analyze all `r nrow(mtcars)` cards in the `mtcars` dataset with the following names:

```{r}
pander(as.list(names(mtcars)))
```

Don't forget to check out the PDF and Word document outputs as well!

