---
title: "A demo R+markdown document"
author: "Gergely Daroczi"
date: '2016-02-03'
output: html_document
---

Markdown is a simple markup language to format plain text files, eg using *italic* or **bold** font, or sometimes even ***both*** or ~~none~~ at all. We can also refer to inline `code` examples, or use syntax highlighting for larger code chunks:

```r
n <- 100
x <- rpois(n, 10)
summary(x)
hist(x)
```

But we can also evaluate `r "R"` code inline or in separate code chunks as well:

```{r}
n <- 100
x <- rpois(n, 10)
summary(x)
hist(x)
```

We can generate markdown tables to be rendered nicely in PDF/HTML/docx etc documents by `kable` from the `knitr` package, or `pander` for the more complex tables:

```{r}
knitr::kable(head(iris))
library(pander)
pander(head(iris))
pander(table(iris$Species))
```

And this latter `pander` package has some nifty features:

```{r}
pander(lm(mpg ~ wt + hp, mtcars))
emphasize.cells(which(head(iris) < 5, arr.ind = TRUE))
pander(head(iris))
```

We can suppress the R commands easily with one of the many chunk options:

```{r, echo = FALSE}
pander(table(iris$Species), justify = 'right')
```

Or we can use global options, eg:

```{r global_options, include = FALSE}
library(knitr)
opts_chunk$set(
    fig.width  = 8,
    fig.height = 8,
    dpi        = 300,
    fig.path   = 'plots/foobar',
    echo       = FALSE,
    warning    = FALSE,
    message    = FALSE)
```
